---
title: "Assignment 1.3"
author: "Saurabh Biswas"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- Import necessary packages -->
```{r, echo=FALSE, message=FALSE}
setwd("/Users/saurabhbiswas/Desktop/DSC630_R")
library(readxl)
library(ggplot2)
```

<!-- Load data from Bureau of Labor Statistics data -->
```{r}
state_M2019_dl <- read_excel("/Users/saurabhbiswas/Desktop/DSC630_R/state_M2019_dl.xlsx")
nebraska_2019 <- subset(state_M2019_dl, state_M2019_dl$area_title == "Nebraska")
```

***
## 1. Import, Plot, Summarize, and Save Data

### **tot_emp:**
```{r}
summary(as.numeric(nebraska_2019$tot_emp))
```

### **h_mean:**	Mean hourly wage 
```{r}
summary(as.numeric(nebraska_2019$a_mean))
```

### **Histogram of occ_code(1-2)**.
```{r}
occ_code_12 <- substr(nebraska_2019$occ_code,1,2)
occ_code_12 <- as.numeric(occ_code_12)
hist(occ_code_12, main ="Histogram of Occupation Code")
```

### **Box plot of hourly mean wage**
```{r}
boxplot(as.numeric(nebraska_2019$h_mean), main='Hourly mean wage boxplot', ylab="Hourly mean Wage")
```

### **Density plot of hourly mean wage**
```{r}
ggplot(nebraska_2019, 
       aes(x = as.numeric(nebraska_2019$h_mean), 
           fill = substr(nebraska_2019$occ_code,1,2))) +
    geom_density(alpha = 0.4) +
    labs(title = 'Salary distribution') +
    xlab("Hourly Mean Salary")
```

### **Density plot of hourly median wage**
```{r}
ggplot(nebraska_2019, 
       aes(x = as.numeric(nebraska_2019$h_mean), 
           fill = o_group)) +
    geom_density(alpha = 0.4) +
    labs(title = 'Salary distribution') +
    xlab("Hourly Median Salary")
```

##### **Write into CSV file**
```{r}
write.csv(nebraska_2019, "nebraska_2019.csv")
```

***

***
## 2. Explore Some Bivariate Relations

### **Scatter plot between annual mean wage and occupation code**
```{r, message=FALSE, warning=FALSE}
plot(as.numeric(substr(nebraska_2019$occ_code,1,2)), as.numeric(nebraska_2019$a_mean),  main="Occupation Code v/s Annual Mean Wage - scatter plot", 
     ylab="annual mean wage", xlab="Occupation Code")
```
***

***
## 3. Organize a Data Report

### **Structure and Type of variables**
```{r}
str(nebraska_2019)
```

### 1. **occ_code** is a chr field. First two positions represent occupation types and next four positions repersent sub-types.

### 2. **tot_emp** is a nueric field. It represents total number of employess for each occupation codes.

### 3. **h_mean** is a float field. It represents hourly mean wage for each occupation codes.

### 4. **h=a_mean** is a float field. It represents annual mean wage for each occupation codes.
***